---
title: "Preparing the Data"
output: html_notebook
---

# Preparing the Data

Data is often stored in a relational way, and instead of having all the variables you need in a single table, you have multiple tables and files that contain different variables.

In this example, we have four different CSV files stored in the *data* directory containing different information about the participants:

+------------------+--------------------------+
| File             | Description              |
+==================+==========================+
| diagnosis.csv    | Mental health conditions |
+------------------+--------------------------+
| groups.csv       | Experimental groups      |
+------------------+--------------------------+
| participants.csv | Demographics             |
+------------------+--------------------------+
| test_results.csv | Experiment Results       |
+------------------+--------------------------+

Before we proceed with our analysis, we'll need to combine the data from the CSV files. The output should be a single tibble with the following columns:

+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| Column                     | Source           | Description                                                                                                    |
+============================+==================+================================================================================================================+
| `participant_id`           | participants.csv | Unique identifier for each partitipant                                                                         |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `birth_date`               | participants.csv | Birth date of each participant                                                                                 |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `age`                      | participants.csv | Age of each participant (range 8-14)                                                                           |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `race_type`                | participants.csv | Race of each participant (White, Hispanic, Mixed/Other)                                                        |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `gender`                   | participants.csv | Male/Female                                                                                                    |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `no_disability`            | diagnosis.csv    | 0: Participant had a disability                                                                                |
|                            |                  |                                                                                                                |
|                            |                  | 1: Participant had no disability                                                                               |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `group`                    | groups.csv       | 1: One-On-One Delivery (ThinkRx)                                                                               |
|                            |                  |                                                                                                                |
|                            |                  | 2: Mixed Delivery (BrainLab)                                                                                   |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `associative_memory_gain`  | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Visual Auditory Learning Subtest     |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `visual_processing_gain`   | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Spatial Relations Subtest            |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `auditory_processing_gain` | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Sound Blending Subtest               |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `logic_reasoning_gain`     | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Concept Formation Subtest            |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `processing_speed_gain`    | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Visual Matching Subtest              |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `working_memory_gain`      | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Numbers Reversed Subtest             |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `long_term_memory_gain`    | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Visual Auditory 1 Learning Subtest   |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+
| `iq_gain`                  | test_results.csv | Improvement in Woodcock III Test of Cognitive Abilities after experiment: Weighted composite of other subtests |
+----------------------------+------------------+----------------------------------------------------------------------------------------------------------------+

## Exercise

```{r start_tidyverse}
library(tidyverse)
```

Start by reading each of the four CSV files into separate tibbles.

```{r read_data_exercise}
# TODO: 
# Load diagnosis.csv into a tibble called `diagnosis_df`
# Load groups.csv into a tibble called `groups_df`
# Load participants.csv into a tibble called `participants_df`
# Load test_results.csv into a tibble called `test_results_df`
# YOUR CODE BELOW
diagnosis_df <- read_csv("../data/diagnosis.csv")
groups_df <- read_csv("../data/groups.csv")
participants_df <- read_csv("../data/participants.csv")
test_results_df <- read_csv("../data/test_results.csv")
```

Take a look at your four tibbles. Notice each row in all of them represent a participant, and each participant has a unique identifier (`participant_id`). This means, we can join the four tibbles into a single tibble using the participant identifiers.

We'll use an outer join (`full_join()`), but it actually doesn't matter in this example because inner, left, or, right joins would yield the same result. All four tibbles contain the same participants.

```{r join_data_exercise}
# TODO: 
# Join the four tibbles into a single tibble called `experiment_data_df`
# Use `participant_id` as the key for the full join
# Inspect the tibble after completing the join
# YOUR CODE BELOW
experiment_data_df <- participants_df %>%
  full_join(diagnosis_df, by = "participant_id") %>%
  full_join(groups_df, by = "participant_id") %>%
  full_join(test_results_df, by = "participant_id")

experiment_data_df
```

Save your results to a CSV file to `../data/experimental_data.csv`. You'll need it for the next exercise.

```{r}
# TODO: Save experiment_data_df to ../data/experimental_data.csv

experiment_data_df %>% write_csv("../data/experimental_data.csv")
```
