---
title: "Validating Random Assignment of Groups"
output: html_notebook
---

# Validating Random Assignment of Groups

We recommend completing and running the code in Load and View Data from CSV Files *(01_prepare_data.Rmd)* before beginning this exercise. Otherwise, run the code chunk below.

```{r source_required_scripts, include=FALSE}
source("../scripts/01_prepare_data.R")
```

## Exercise

In this notebook, you will validate that groups were randomly assigned to individuals by checking for differences in demographics between the groups.

### Age

First, we'll test whether the difference between the mean ages in the one-on-one and mixed delivery experimental groups are different.

|                        |                                                                                                       |
|--------------------|----------------------------------------------------|
| Null Hypothesis        | The difference in means of ages between one-on-one delivery and mixed delivery is equal to 0.         |
| Alternative Hypothesis | The difference in means of ages between one-on-one delivery and mixed delivery is **not** equal to 0. |

For this, you'll use a [Two Sample t-test](https://www.statology.org/two-sample-t-test/). Below is the code to run the test in R:

    t.test(
      experiment_data_df$age ~ experiment_data_df$group, # Testing age as a function of group
      alternative = "two.sided",                         # Two-sided test ("not equal to 0")
      var.equal = TRUE                                   # Treat variance of two groups as equal
    )

Go ahead and try running the code yourself.

```{r check_random_age}
# TODO: 
# Complete Two Sample t-test for age as a function of group
# YOUR CODE BELOW
t.test(
  experiment_data_df$age ~ experiment_data_df$group,
  alternative = "two.sided",
  var.equal = TRUE
)
```

Based on the p-value, we don't have enough evidence to reject the null hypothesis at the 5 percent significance level. Therefore, we accept the null hypothesis:

|                                                                                               |
|------------------------------------------------------------------------|
| The difference in means of ages between one-on-one delivery and mixed delivery is equal to 0. |

This mean we conclude that there is no meaningful difference in ages between the two experimental groups.

### Gender

*\*Generally, you should check whether the term "Gender" is being used synonymously with "Sex Assigned at Birth." This is especially important when male and female are the only responses.*

Next, let's test whether the the experimental groups and gender are independent.

|                        |                                                        |
|--------------------|----------------------------------------------------|
| Null Hypothesis        | There is **no** relationship between gender and group. |
| Alternative Hypothesis | There is relationship between gender and group.        |

For this, you'll use a Pearson's Chi Squared Test. Below is the code to run the test in R:

    chisq.test(
      experiment_data_df$group,     # Testing whether gender and group are independent
      experiment_data_df$gender,
      correct = FALSE,              # Continuity correction is not applied
      simulate.p.value = TRUE       # p values are computed by Monte Carlo simulation
    )

More information about:

-   [Pearson's Chi Squared Test](https://en.wikipedia.org/wiki/Pearson%27s_chi-squared_test)

-   [Continuity Correction](https://en.wikipedia.org/wiki/Yates%27s_correction_for_continuity)

-   [Monte Carlo Simulation](https://stats.stackexchange.com/questions/62445/rules-to-apply-monte-carlo-simulation-of-p-values-for-chi-squared-test)

Go ahead and try running the code yourself.

```{r check_random_gender}
# TODO: 
# Complete Chi Squared test for group and gender
# YOUR CODE BELOW
chisq.test(
  experiment_data_df$group,     # Testing whether gender and group are independent
  experiment_data_df$gender,
  correct = FALSE,              # Continuity correction is not applied
  simulate.p.value = TRUE       # p values are computed by Monte Carlo simulation
)
```

Based on the p-value, we don't have enough evidence to reject the null hypothesis at the 5 percent significance level. Therefore, we accept the null hypothesis:

|                                                        |
|--------------------------------------------------------|
| There is **no** relationship between gender and group. |

This mean we conclude that there is no meaningful relationship between gender and the experimental groups.

### Race and Disability Status

Both race and disability status are categorical data like gender, and we should use the chi-squared test to determine whether they are independent from experimental groups. Like before, we will not apply the Continuity Correction, and we will use Monte Carlo simulations to calculate the p-value.

```{r check_random_race}
# TODO: 
# Complete Chi Squared test for group and race
# YOUR CODE BELOW
chisq.test(
  experiment_data_df$group,     # Testing whether gender and group are independent
  experiment_data_df$race_type,
  correct = FALSE,              # Continuity correction is not applied
  simulate.p.value = TRUE       # p values are computed by Monte Carlo simulation
)
```

```{r check_random_disability}
# TODO: 
# Complete Chi Squared test for group and disability status
# YOUR CODE BELOW
chisq.test(
  experiment_data_df$group,     # Testing whether gender and group are independent
  experiment_data_df$no_disability,
  correct = FALSE,              # Continuity correction is not applied
  simulate.p.value = TRUE       # p values are computed by Monte Carlo simulation
)
```

Verify that your results match the results from the [research paper](https://rdcu.be/c3NWc), which can be found in the *Procedures* section on page 263:

*"A check of the random assignment indicated the groups were balanced, with no significant differences between groups based on personal characteristics (age: t = 0.217, p = 0.829; gender: χ 2 = 1.80, p = 0.210; race/ethnicity: χ2 = 2.35, p = 0.218; no disability: χ2 = 0.54, p = 0.522)."*
