---
title: "Testing for Significant Differences in Response Variables"
output: html_notebook
---

# Testing for Significant Differences in Response Variables

We recommend completing and running the code in Summarize Demographics *(01_summarize_demographics.Rmd)* before beginning this exercise. Otherwise, run the code chunk below.

```{r source_required_scripts, include=FALSE}
source("../scripts/01_summarize_demographics.R")
```

In addition, you should review the *Results* section of the [research paper](https://rdcu.be/c3NWc) before beginning. Pay close attention to *4.2. The Impact of Layout, Benefit Type and Heritage on Buying Intention*.

In this notebook, we'll reproduce the results in *4.2. The Impact of Layout, Benefit Type and Heritage on Buying Intention*.

## Exercise

We'll use a [3-Way ANOVA test](https://www.statology.org/three-way-anova/). To run a 3-Way ANOVA test in R, you'll use the `aov()` and `summary.aov()` functions. Look up the documentation for the `aov()` and `summary.aov()` function, and perform ANOVA with buying intention as a function of the layout, benefit, and heritage independent variables.

```{r test_anova}
# TODO: 
# Complete ANOVA test with buying intention as a function of the layout, benefit, and heritage independent variables
# Display a summary of the results
# YOUR CODE BELOW
anova_test <- aov(
  BI_rec ~ IV1_lay * IV2_ben * IV3_her,           
  data = study_data_df
)

summary.aov(anova_test)
```

Which main effect and interactions were significant at the 10 percent significance level? Your results should match those described in the paper.
