---
title: "Examine Data"
output: html_notebook
---

# Examine Subsets of Data

In this notebook, you will examine subsets of the data. You will first review an example and then try it yourself by completing an exercise.

## Supporting Materials

Make sure you are familiar with the knowledge covered in:

-   R for Everyone
-   R for Data Manipulation

In addition, we recommend completing and running the code in the following exercises before beginning this one:

-   *Load and View Data from CSV Files* *(01_load_data.Rmd)*

-   *Concatenate Data from Multiple Tibbles* *(02_concatenate_data.Rmd)*

Otherwise, run the code chunk below.

```{r source_required_scripts, include=FALSE}
source("../scripts/02_concatenate_data.R")
```

## Example

Suppose we are interested in the different between the post assessment and pre assessment scores. To do this, we'll first create a column called *assessment_progress*.

```{r mutate_example}
class_attendance_df %>%
  mutate(assessment_progress = Post.Assessment.Score - Pre.Assessment.Score)
```

Let's first take a look at *assessment_progress* for class a. To do this, we'll use the `filter()` function to filter to students of class a and the `select()` function to view just the *assessment_progress* column.

```{r }
class_attendance_df %>%
  mutate(assessment_progress = Post.Assessment.Score - Pre.Assessment.Score) %>%
  filter(class == "a") %>%
  select(assessment_progress)
```

It looks like the majority of students in class a displayed progress (*assessment_progress* \> 0).

Now let's suppose we wanted to compare *assessment_progress* across the three classes. Specially, let's compare the number of students, the average assessment progress, and the median assessment progress. To do this, we'll need to:

-   Use `mutate()` to create the *assessment_progress* column

-   Use `group_by()` to specify that you would like to compare across the *class* column

-   Use `summarize()` to indicate that we're looking for the number of students (`n()`), the average assessment progress (`mean()`), and the median assessment_progress(`median()`)

```{r summarize_example}
class_attendance_df %>%
  mutate(assessment_progress = Post.Assessment.Score - Pre.Assessment.Score) %>%
  group_by(class) %>%
  summarize(
    students = n(),
    average_change = mean(assessment_progress),
    median_change = median(assessment_progress)
  )
```

## Exercise

Now it's your turn. In addition to assessment progress, let's say you're interested in there total attendance (total number of classes attended) of each student. Here is your task for this exercise:

1.  Create a new column in *class_attendance_df* called *total_attendance*
2.  Summarize the total students, average total attendance, and median total attendance by class

```{r summarize_exercise}
# TODO: Create a new column in class_attendance_df called total_attendance
# TODO: Summarize the total students, average total attendance, and median total attendance by class
# YOUR CODE BELOW
class_attendance_df %>%
  mutate(total_attendance = Week.1.Attendance + Week.2.Attendance + Week.3.Attendance + Week.4.Attendance + Week.5.Attendance) %>%
  group_by(class) %>%
  summarize(
    students = n(),
    average_total_attendance = mean(total_attendance),
    median_total_attendance = median(total_attendance)
  )
```
