---
title: "Challenge"
output: html_notebook
---

# Class Attendance

## Supporting Materials

Make sure you are familiar with the knowledge covered in:

-   R for Everyone
-   R for Data Manipulation

In addition, we recommend completing the following exercises before beginning the challenge:

-   *Load and View Data from CSV Files* *(01_load_data.Rmd)*

-   *Concatenate Data from Multiple Tibbles* *(02_concatenate_data.Rmd)*

-   *Examine Subsets of Data (03_examine_data.Rmd)*

## Challenge

We have six classes (a, b, c, d, e, and f) all taking the same course, and we would like to know whether older students miss fewer classes and show more progress.

Your task is to load the csv files for all six classes and create a single table that displays the following columns:

| Column                      | Description                                                                                               |
|:------------------------|:---------------------------------------------|
| Grade                       | A grade of students (6, 7, or 8)                                                                          |
| students                    | The number of students in the grade (across all six classes)                                              |
| average_total_attendance    | The class mean for the total attendance of each student                                                   |
| median_total_attendance     | The class median for the total attendance of each student                                                 |
| average_assessment_progress | The class mean for the assessment progress (difference between post and pre assessment) of each student   |
| median_assessment_progress  | The class median for the assessment progress (difference between post and pre assessment) of each student |

```{r}
library(tidyverse)
```

```{r}
class_attendance_df <- c()
for (class_csv in list.files("../data")){
  print(class_csv)
  class_attendance_df <- bind_rows(
    class_attendance_df,
    read_csv(paste("../data/", class_csv, sep = "")) %>% 
      mutate(class = substr(class_csv, 7, 7)),
  )
}

class_attendance_df
```

```{r}
class_attendance_df %>%
  mutate(
    total_attendance = Week.1.Attendance + Week.2.Attendance + Week.3.Attendance + Week.4.Attendance + Week.5.Attendance,
    assessment_progress = Post.Assessment.Score - Pre.Assessment.Score
  ) %>%
  group_by(Grade) %>%
  summarize(
    students = n(),
    average_total_attendance = mean(total_attendance),
    median_total_attendance = median(total_attendance),
    average_assessment_progress = mean(assessment_progress),
    median_assessment_progress = median(assessment_progress)
  )
```
