---
title: "Load and View Data from CSV files"
output: html_notebook
---

# Load and View Data from CSV Files

## Supporting Materials

## Example

First we need to load the libraries we're using\*. In this example, all we're using is the `tidyverse` library.

*\*You'll need to ensure that your libraries are installed (`install.packages()`) first before loading them.*

```{r load_required_libraries}
library(tidyverse)
```

Now let's load in the data we're analyzing. In this example, we'll load in *class_a.csv*.

While loading in data from CSV files, it's important to understand the location of the CSV files relative to the location of the R Notebook. Take a look at the diagram below.

    class_attendance/
    ├── notebooks/
    │   ├── 01_load_data.Rmd
    │   └── 01_merge_data.Rmd
    └── data
        ├── class_a.csv
        └── class_b.csv

The notebook we're using right now is *01_load_data.Rmd* and the CSV file we're trying to load is *class_a.csv*. Notice that they're in separate folders (*notebooks* and *data*). To reach *data/class_a.csv* from *notebooks/01_load_data.Rmd*, we need to:

1.  Step out of the *notebooks* directory

2.  Step into the *data* directory

3.  Access *class_a.csv*

You can think of this as the **path** from *notebooks/01_load_data.Rmd* to *data/class_a.csv* and here is how it's represented.

+----------------------------------------------------------------------------+-----------------------+
| Step                                                                       | Representation        |
+============================================================================+=======================+
| Full relative path from *notebooks/01_load_data.Rmd* to *data/class_a.csv* | `../data/class_a.csv` |
+----------------------------------------------------------------------------+-----------------------+
| 1.  Step out of the *notebooks* directory                                  | `../`                 |
+----------------------------------------------------------------------------+-----------------------+
| 2.  Step into the *data* directory                                         | `data/`               |
+----------------------------------------------------------------------------+-----------------------+
| 3.  Access *class_a.csv*                                                   | `class_a.csv`         |
+----------------------------------------------------------------------------+-----------------------+

Now that you have the path (`../data/class_a.csv`), you can use the `read_csv()` function to load the CSV file into a tibble.

```{r read_data_a}
class_a_df <- read_csv("../data/class_a.csv")
```

With the data now stored in a tibble called `class_a_df`, we can now view the data in our notebook.

```{r view_data_a}
class_a_df
```

## Exercise
